### 2018-10-17 Latex/section_methods.tex
line length

We have carried out a series of twenty forced-dissipative simulations of the SW equations in a quadratic domain with periodic boundary conditions using a standard pseudo spectral code. The simulations were started from zero initial conditions and were run till a statistically stationary state was reached. The equations 
were reformulated in terms of  the three eigenmodes of the linearized equations in Fourier space. One of these eigenmodes is zero for an irrotational flow.  
We first carried out a few test simulations in which all three eigenmodes were allowed to evolve in time. However, when only the two wave modes are forced vorticity remains very small throughout the simulations and does not affect the wave dynamics. Therefore, we chose to solve the equations only for the two wave modes. As discussed in the previous section,  a linear viscous term in (\ref{eq_uu}) 
will not conserve momentum exactly and does not give a strictly positive definite expression for energy dissipation. We have relaxed the requirements of exact momentum conservation and positive definite energy dissipation. 
Instead of  using the nonlinear form of the viscous term in (\ref{eq_uu}) we have used a higher order linear operator  $ \nu_8 \nabla^{8} $ both in (\ref{eq_uu}) and
(\ref{eq_h}), where $ \nu_8 $ is a `hyperdiffusion' or a `hyperviscosity'. There is no strong physical motivation for this
modification but, as discussed by \cite{FargeSadourny1989}, a molecular viscous operator 
 of the form $\nu \nabla^2 $ is not necessarily a
relevant model of the actual small-scale dissipation for SW
flows, which should rather be described as a transition from
two-dimensional to three-dimensional motions before reaching the
scales where dissipation actually occurs.

@@

Time advancement is carried out by a classical fourth-order
Runge-Kutta scheme for the nonlinear term and an exact integration for
the linear and dissipative terms.  This explicit integration is
especially useful at very high resolution and very large $c$ when
the shortest waves are very fast. 
We use an adaptable time step method which maximizes the time step
over a standard Courant-Friedrichs-Lewy condition
\cite[]{Lundbladh1999, AugierChomazBillant2012}.

@@

Most of the aliasing is removed by truncating 8/9 of the modes along
each direction \cite[for a detail discussion on the issues of the
non-conservation of the non-quadratic energy and the aliasing errors
in the truncated one-layer shallow water model,
see][]{FargeSadourny1989}. The resolution is characterized by the number of nodes, $n$, in the each
direction and has been varied from 960 to 7680.

@@

The wave speed $c$ has been varied over almost two orders of magnitude from
10 to 700. The wave modes are forced in a shell in spectral space 
corresponding to relatively small wave numbers, 
$ 5 \delta k <  |\kk| <  8 \delta k$,  
where  $ \delta k = 2 \pi /L $ and $ L $ is the length of the side of the quadratic domain. 
The forcing is white noice in time and designed so that the energy injection rate is normalized to unity at each time step.
The energy is dissipated at the smallest resolved scales by hyperviscosity.
The value of the
hyperviscosity, $\nu_8$, is chosen in such a way that the  maximum wave number is larger than the dissipation wave number

@@

 where $ \eps $ is the mean energy dissipation rate. We found that  $
 k_{d} \simeq \kmax/ 2.5$, where $\kmax =
(8/9)\pi n/L $ is the maximum wave number, is  sufficient in order for dissipation spectra to be resolved. Since the energy injection rate is normalized to unity and all simulations reach a stationary state in which dissipation is equal to the injection rate, $ \nu_8 $ can be chosen already from the start so that this condition will be fulfilled in the stationary state.  The strength of the stratification is characterized by a Froude number, defined as

@@

\begin{figure}
 \centerline{
 \includegraphics[width=3.15in]{../Pyfig/fig_spatiotempspectra}}
\caption{Frequency spectra of KE  and APE for different $ k = \mid {\bf k} \mid $, from run W7. The spectra are plotted  as functions of $\omega/\omega_l$, where
$\omega_l = c k$.  From top to bottom: $ k /\delta k
= 12$, 27, 62, 143, 327, and 746.  }
\label{fig_spatiotemp_spectra}
\end{figure}
\begin{figure}
 \centerline{
 \includegraphics[width=6.0in]{../Pyfig/fig_phys_fields_wave}}
\caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
\label{Physical}
\end{figure}
\begin{figure}
\centerline{\includegraphics[width=8cm]{../Pyfig/fig_shock_sep.pdf}}
\caption{Mean distance between the shocks as function of Froude number.  }
\label{fig_distance}
\end{figure}

>>>>

We have carried out a series of twenty forced-dissipative simulations of the SW
equations in a quadratic domain with periodic boundary conditions using a
standard pseudo spectral code. The simulations were started from zero initial
conditions and were run till a statistically stationary state was reached. The
equations were reformulated in terms of the three eigenmodes of the linearized
equations in Fourier space. One of these eigenmodes is zero for an irrotational
flow. We first carried out a few test simulations in which all three eigenmodes
were allowed to evolve in time. However, when only the two wave modes are
forced vorticity remains very small throughout the simulations and does not
affect the wave dynamics. Therefore, we chose to solve the equations only for
the two wave modes. As discussed in the previous section, a linear viscous term
in (\ref{eq_uu}) will not conserve momentum exactly and does not give a
strictly positive definite expression for energy dissipation. We have relaxed
the requirements of exact momentum conservation and positive definite energy
dissipation. Instead of using the nonlinear form of the viscous term in
(\ref{eq_uu}) we have used a higher order linear operator $ \nu_8 \nabla^{8} $
both in (\ref{eq_uu}) and (\ref{eq_h}), where $ \nu_8 $ is a `hyperdiffusion'
or a `hyperviscosity'. There is no strong physical motivation for this
modification but, as discussed by \cite{FargeSadourny1989}, a molecular viscous
operator of the form $\nu \nabla^2 $ is not necessarily a relevant model of the
actual small-scale dissipation for SW flows, which should rather be described
as a transition from two-dimensional to three-dimensional motions before
reaching the scales where dissipation actually occurs.

@@

Time advancement is carried out by a classical fourth-order Runge-Kutta scheme
for the nonlinear term and an exact integration for the linear and dissipative
terms. This explicit integration is especially useful at very high resolution
and very large $c$ when the shortest waves are very fast. We use an adaptable
time step method which maximizes the time step over a standard
Courant-Friedrichs-Lewy condition \cite[]{Lundbladh1999,
AugierChomazBillant2012}.

@@

Most of the aliasing is removed by truncating 8/9 of the modes along each
direction \cite[for a detail discussion on the issues of the non-conservation
of the non-quadratic energy and the aliasing errors in the truncated one-layer
shallow water model, see][]{FargeSadourny1989}. The resolution is characterized
by the number of nodes, $n$, in the each direction and has been varied from 960
to 7680.

@@

The wave speed $c$ has been varied over almost two orders of magnitude from 10
to 700. The wave modes are forced in a shell in spectral space corresponding to
relatively small wave numbers, $ 5 \delta k < |\kk| < 8 \delta k$, where $
\delta k = 2 \pi /L $ and $ L $ is the length of the side of the quadratic
domain. The forcing is white noice in time and designed so that the energy
injection rate is normalized to unity at each time step. The energy is
dissipated at the smallest resolved scales by hyperviscosity. The value of the
hyperviscosity, $\nu_8$, is chosen in such a way that the maximum wave number
is larger than the dissipation wave number

@@

where $ \eps $ is the mean energy dissipation rate. We found that $ k_{d}
\simeq \kmax/ 2.5$, where $\kmax = (8/9)\pi n/L $ is the maximum wave number,
is sufficient in order for dissipation spectra to be resolved. Since the energy
injection rate is normalized to unity and all simulations reach a stationary
state in which dissipation is equal to the injection rate, $ \nu_8 $ can be
chosen already from the start so that this condition will be fulfilled in the
stationary state. The strength of the stratification is characterized by a
Froude number, defined as

@@

\begin{figure}
\centerline{\includegraphics[width=3.15in]{../Pyfig/fig_spatiotempspectra}}
\caption{Frequency spectra of KE and APE for different $ k = \mid {\bf k} \mid
$, from run W7. The spectra are plotted as functions of $\omega/\omega_l$,
where $\omega_l = c k$. From top to bottom: $ k /\delta k = 12$, 27, 62, 143,
327, and 746. }
\label{fig_spatiotemp_spectra}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=6.0in]{../Pyfig/fig_phys_fields_wave}}
\caption{Left: divergence $ \bnabla \cdot {\bf u} $. Right: velocity component in the $ y $-direction: $ u_y $. Top: run W6. Bottom: run W17.  }
\label{Physical}
\end{figure}

\begin{figure}
\centerline{\includegraphics[width=8cm]{../Pyfig/fig_shock_sep.pdf}}
\caption{Mean distance between the shocks as function of Froude number.  }
\label{fig_distance}
\end{figure}
